---
title: "workshop-scraping (solutions)"
author: "Maël Lecoursonnais"
date: "2023-06-09"
output: html_document
---

In this lab, we are going to see how to handle spatial data using OpenStreetMap.

OpenStreetMap (OSM) is a free, open geographic database updated and maintained by a community of volunteers via open collaboration [1](https://en.wikipedia.org/wiki/OpenStreetMap).

It provides an easily accessible API from which we can retrieve spatial data. In R, we can use the `osmdata` package to get tidy output from the API query.

## Prerequisites

```r
install.packages("tidyverse")
install.packages("osmdata")
install.packages("sf")
install.packages("tmap")
install.packages("leaflet")
```

## Retrieving data on OSM

With `osmdata`, a query can be done like this:

  1. Define an area from which you want to collect spatial information. This area is called a **bounding box**. Use `opq` and/or `getbb`. 
  
  2. Add the features you want to collect. A list can be found [here](https://wiki.openstreetmap.org/wiki/Map_features) (or use `osmdata::available_features()`). A feature consists of a key (a broad category, e.g. a building) and a value (e.g., a church). Use `add_osm_feature`.
  
  3. Convert the result to a `sf` object. Use `osmdata_sf`.

For instance, let's find out the schools of Norrköping:
```{r include=FALSE}
library(tidyverse)
library(osmdata)
library(sf)

nkpg_schools <- 
  #Create the query and set the bounding box
  opq(bbox = "Norrköping") %>% 
  #Select the features
  add_osm_feature(key = "amenity", value = "school") %>%
  #Convert to osmdata_sf
  osmdata_sf()
```

OSM data can be a bit messy, since this is a collaborative effort. Here, schools are sometimes points, lines, polygons, or multipolygons.

## Mapping data

It is a bit uncertain as to what we captured with this query: are the schools really in Norrköping? Are they really schools? It's always better to double-check with a map! Hello.

Here, there are two options: static or interactive maps. Several packages allow to map `sf` objects. We are going to focus on `ggplot2` for static maps and `tmap` for interactive ones. `leaflet` is also a good options for interactive maps.

```{r}
library(tmap)
tmap_mode("view") #Set interactive mode

#Create a data.frame with all schools. Careful, this data.frame includes multiple geometries!
nkpg_schools_sf <- bind_rows(
  st_make_valid(nkpg_schools$osm_multipolygons),
  nkpg_schools$osm_polygons[!is.na(nkpg_schools$osm_polygons$name), ],
  nkpg_schools$osm_points[!is.na(nkpg_schools$osm_points$name), ]
  ) 

#Map
tm_basemap("Esri.WorldGrayCanvas") +
  tm_shape(nkpg_schools_sf) +
  tm_polygons(popup.vars = "name") +
  tm_shape(nkpg_schools_sf[st_dimension(nkpg_schools_sf) == 0, ]) +
  tm_dots(popup.vars = "name")
```

## Exercise 1

  1. Using the pipeline described above, find all the hospitals in Stockholm.
  2. Map the hospitals.
  
```{r}
sklm_hospitals <- 
  #Create the query and set the bounding box
  opq(bbox = "Stockholm") %>% 
  #Select the features
  add_osm_feature(key = "amenity", value = "hospital") %>%
  #Convert to osmdata_sf
  osmdata_sf()

sklm_hospitals_sf <- bind_rows(
  st_make_valid(sklm_hospitals$osm_multipolygons),
  sklm_hospitals$osm_polygons[!is.na(sklm_hospitals$osm_polygons$name), ],
  sklm_hospitals$osm_points[!is.na(sklm_hospitals$osm_points$name), ]
  ) 

tm_basemap("Esri.WorldGrayCanvas") +
  tm_shape(sklm_hospitals_sf) +
  tm_polygons(popup.vars = "name") +
  tm_shape(sklm_hospitals_sf[st_dimension(sklm_hospitals_sf) == 0, ]) +
  tm_dots(popup.vars = "name")
```

## Exercise 2

Using `osmdata::getbb`, get the boundaries of the city of Paris, and of Sweden. Check `?getbb`.

```{r}
#Paris 
paris <- getbb("Paris, France", format_out = 'sf_polygon', featuretype = "settlement")
plot(st_geometry(paris))

#Sweden
sweden <- getbb("Sweden", format_out = 'sf_polygon', featuretype = "country")
plot(sweden$multipolygon)
```

  










